---
title: 'Neil Shah: DATA 607 HW 9'
output:
  html_document:
    df_print: paged
---

## Intrdouction

The purpose of this notebook is to discuss how to develop a function that can interface with the New York Times API. After reading the [documentation](https://developer.nytimes.com/apis), which is critical to understand any API, I registered for an unique API key and planned on making a dataframe using the described JSON architecture.

## Packages Involved

```{r}
library(tidyverse)
library(jsonlite)
library(DT)
library(RCurl)
library(stringr)
```

## API Use

The NY Times has 12 different APIs as follows:

1. Archive 
2. Article Search
3. Books
4. Community (beta)
5. Geo
6. Most Popular
7. Movie Reviews
8. RSS Feeds
9. Semantic 
10. Times Tag
11. Timews Wire
12. Top Stories

I chose to work with the Articles API since I'm news junky!

First  saved my api key to a variable

```{r}
api = "gGojTtfdRkDfGOJxPfPi7ZeLAr6O4Wri"
```

Now looking at the API details the call formula is relatively simple and takes the form of 

```
https://api.nytimes.com/svc/topstories/v2/CATEGORY.json?api-key
```

Where category can be one of the following:

arts
automobiles
books
business
fashion
food
health
home
insider
magazine
movies
nyregion
obituaries
opinion
politics
realestate
science
sports
sundayreview
technology
theater
t-magazine
travel
upshot
us
world.

Let's do an example using 'arts' as a category. 

We already know that the API will spit out a JSON so we can use fromJSON to read it.

We will need to stitch the url base and the api together--so let's use paste0 (which has no space unlike space)

```{r}
head(fromJSON(paste0('https://api.nytimes.com/svc/topstories/v2/arts.json?api-key=',api)))
```


Ok look's simple enough--let's throw it into a dataframe and a tibble

```{r}
fromJSON(paste0('https://api.nytimes.com/svc/topstories/v2/arts.json?api-key=',api))  %>% data.frame() %>% as.tibble() %>% head()
```


Well that was easy...the End?

  
## The End?
  
Nope! We both know that Layla is doing something exemplary--and I can't sleep on that! Lets make a handy function that we can specify a category and it returns our dataframe!
  
  First let's make a vector of subjects.

```{r}
subjects= c('arts','automobiles','books','business','fashion','food','health','home','insider','magazine','movies','nyregion','obituaries','opinion','politics','realestate','science','sports','sundayreview','technology','theather','travel','upshot','us','world')
```

So now it should be a simple as passing a category to our function, pasting it in a URL and then making a dataframe.

Let's do that!

```{r}
NYTarticle <- function(subject='home')
{
#This function returns a dataframe of NYTarticles--it defaults subjects to home and api to my key
#First let's make a check that a valid subject is passed
subjectslist= c('arts','automobiles','books','business','fashion','food','health','home','insider','magazine','movies','nyregion','obituaries','opinion','politics','realestate','science','sports','sundayreview','technology','theather','travel','upshot','us','world')
  if (typeof(subject)!='character')
  {
    print('Subject must be a charcter type!')
  }
  else if (as.character(subject) %in% subjectslist)
  {
    articledf <- fromJSON(paste0('https://api.nytimes.com/svc/topstories/v2/',subject,'.json?api-key=',"gGojTtfdRkDfGOJxPfPi7ZeLAr6O4Wri")) %>% data.frame() %>% as_tibble()
    return(articledf)
  }
  else
  {
    print('Invalid Subject: Please select from following list or  leave blank for home')
    print(subjectslist)
  }
  

  
}
```


So what's going on over the hood is 

1. We can pass a subject (defaulted to home) to our function NYTarticle
2. We check to make sure the subject is a character otherwise we stop
3. If it's a character we then check to make sure it's a valid subject by comparing to our list--if it isn't we stop but we give the user the list of options.
4. If it's a valid list, we stitch it with the query URL, pass it to a JSON and return the dataframe!


Let's test it with a non string

```{r}
NYTarticle(5)
```

Now with a subject not in the list
```{r}
NYTarticle('cars')
```

Finally with something in the list
```{r}
NYTarticle('science')
```


## Conclusion

In thie notebook we made a easy to use function that can accept subject fields and return a dataframe of articles--great for coronavirus reading!

